---
title: "Figures"
author: "Anne Baranger"
date: "2024-07-05"
output:
  bookdown::html_document2:
    toc: true            # Table des matières (menu)
    toc_float: true       # Menu flottant sur le côté
    number_sections: true # Numérotation automatique des sections
    code_folding: hide    # Cacher le code par défaut (cliquable pour afficher)
    df_print: paged       # Jolie impression des dataframes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE
)

lapply(c("targets", "stringr", "tidyr", "dplyr", "tibble", # data analysis
         "ggplot2", "viridis", "ggridges",                # plot
         "rstan", "shinystan", "BIOMASS",                 # stat & models
         "rworldmap"                                      # geographic data
         ), require, character.only = TRUE)
```

# Trajectories with systori model

## With global fits concatenation - by origin type

```{r,fig.height=9,fig.width=7}
traj<-tar_read(global_fit_systori)$trajectory
fit<-tar_read(global_fit_systori)$fit_df
tar_load(mod.data)
tar_load(correspondance_table)
trajectories<-traj %>% 
  mutate(system=factor(system,levels=c("forest","old_sf","young_sf","plantation","agroforestry"))) %>% 
  ggplot()+
  geom_point(data=mod.data,aes(dbh,H,color=system),alpha=0.1,guide='none')+
  geom_ribbon(aes(dbh,ymin=H_q05,ymax=H_q95,group=system),fill="grey",alpha=0.5)+
  geom_line(aes(dbh,H_med,group=system,color=system))+
  facet_wrap(~origin)+
  scale_color_manual(labels=c("Forest","Old sec. for.","Young sec. for.","Plantation","Agroforestry"),
                     values=c("darkgreen","mediumseagreen","mediumspringgreen","steelblue3","lightsalmon"))+
  scale_y_log10()+
  theme_light()+
  theme(legend.position = "top")+
  labs(color="System")

alpha<-fit |> 
  select(matches("alpha")&!matches("_0")) |> 
  pivot_longer(cols=everything()) |> 
  mutate(name=str_replace(name,"alpha_",""))|>
  left_join(correspondance_table,by=c("name"="corr")) |> 
  mutate(system=factor(system,levels=c("forest","old_sf","young_sf","plantation","agroforestry"))) %>% 
  ggplot(aes(value,system,fill=system))+
  geom_density_ridges(scale=0.8)+
  scale_fill_manual(labels=c("Forest","Old sec. for.","Young sec. for.","Plantation","Agroforestry"),
                     values=c("darkgreen","mediumseagreen","mediumspringgreen","steelblue3","lightsalmon"))+
  facet_wrap(~origin)+
  theme_light()+
  theme(axis.text.y = element_text(color="white",angle=90),
        axis.ticks.y = element_line(color="white"),
        legend.position = "none")+
  labs(y="",x="Alpha value")

beta<-fit |> 
  select(matches("beta")&!matches("_0")) |> 
  pivot_longer(cols=everything()) |> 
  mutate(name=str_replace(name,"beta_","")) |> 
  left_join(correspondance_table,by=c("name"="corr")) |> 
  mutate(system=factor(system,levels=c("forest","old_sf","young_sf","plantation","agroforestry"))) %>% 
  ggplot(aes(value,system,fill=system))+
  geom_density_ridges(scale=0.8)+
  scale_fill_manual(labels=c("Forest","Old sec. for.","Young sec. for.","Plantation","Agroforestry"),
                     values=c("darkgreen","mediumseagreen","mediumspringgreen","steelblue3","lightsalmon"))+
  facet_wrap(~origin)+
  theme_light()+
  theme(axis.text.y = element_text(color="white",angle=90),
        axis.ticks.y = element_line(color="white"),
        legend.position = "none")+
  labs(y="",x="Beta value")
cowplot::plot_grid(trajectories,alpha,beta,nrow=3,rel_heights = c(1,0.6,0.6))
```

## With global fits concatenation - by system type

```{r,fig.height=9,fig.width=8}
traj<-tar_read(global_fit_systori)$trajectory
fit<-tar_read(global_fit_systori)$fit_df
tar_load(mod.data)
tar_load(correspondance_table)
trajectories<-traj %>% 
  mutate(system=factor(system,levels=c("forest","old_sf","young_sf","plantation","agroforestry")),
         origin=factor(origin,levels=c("planted","recruited","remnant"))) %>% 
  ggplot()+
  geom_point(data=mod.data,aes(dbh,H,color=origin),alpha=0.1,guide='none')+
  geom_ribbon(aes(dbh,ymin=H_q05,ymax=H_q95,group=origin),fill="grey",alpha=0.5)+
  geom_line(aes(dbh,H_med,group=origin,color=origin))+
  facet_wrap(~system,nrow=1)+
  scale_color_manual(labels=c("Planted","Recruited","Remnant"),
                     values=c("darkred","steelblue2","forestgreen"))+
  scale_y_log10()+
  theme_light()+
  theme(legend.position = "top")+
  labs(color="Origin")

alpha<-fit |> 
  select(matches("alpha")&!matches("_0")) |> 
  pivot_longer(cols=everything()) |> 
  mutate(name=str_replace(name,"alpha_",""))|>
  left_join(correspondance_table,by=c("name"="corr")) |> 
  mutate(system=factor(system,levels=c("forest","old_sf","young_sf","plantation","agroforestry")),
         origin=factor(origin,levels=c("planted","recruited","remnant"))) %>% 
  ggplot(aes(value,origin,fill=origin))+
  geom_density_ridges(scale=0.8)+
  scale_fill_manual(labels=c("Planted","Recruited","Remnant"),
                     values=c("darkred","steelblue2","forestgreen"))+
  facet_wrap(~system,nrow = 1)+
  theme_light()+
  theme(axis.text.y = element_text(color="white",angle=90),
        axis.ticks.y = element_line(color="white"),
        legend.position = "none")+
  labs(y="",x="Alpha value")

beta<-fit |> 
  select(matches("beta")&!matches("_0")) |> 
  pivot_longer(cols=everything()) |> 
  mutate(name=str_replace(name,"beta_","")) |> 
  left_join(correspondance_table,by=c("name"="corr")) |> 
  mutate(system=factor(system,levels=c("forest","old_sf","young_sf","plantation","agroforestry")),
         origin=factor(origin,levels=c("planted","recruited","remnant"))) %>% 
  ggplot(aes(value,origin,fill=origin))+
  geom_density_ridges(scale=0.8)+
  scale_fill_manual(labels=c("Planted","Recruited","Remnant"),
                     values=c("darkred","steelblue2","forestgreen"))+
  facet_wrap(~system,nrow=1)+
  theme_light()+
  theme(axis.text.y = element_text(color="white",angle=90),
        axis.ticks.y = element_line(color="white"),
        legend.position = "none")+
  labs(y="",x="Beta value")
cowplot::plot_grid(trajectories,alpha,beta,nrow=3,rel_heights = c(1,0.6,0.6))
```

# comparison of fits

## sample 100
```{r}
tar_load(data.strat.100)
tar_load(data.strat.50)
tar_load(mod.data)

tar_read(global_fit_systori)$trajectory %>% 
  mutate(system=factor(system,levels=c("forest","old_sf","young_sf","plantation","agroforestry")),
         origin=factor(origin,levels=c("planted","recruited","remnant"))) %>% 
  ggplot()+
  geom_point(data=mod.data,aes(dbh,H,color=origin),alpha=0.1,guide='none')+
  geom_ribbon(aes(dbh,ymin=H_q05,ymax=H_q95,group=origin),fill="grey",alpha=0.5)+
  geom_line(aes(dbh,H_med,group=origin,color=origin))+
  facet_wrap(~system,nrow=1)+
  scale_color_manual(labels=c("Planted","Recruited","Remnant"),
                     values=c("darkred","steelblue2","forestgreen"))+
  scale_y_log10()+
  theme_light()+
  theme(legend.position = "top")+
  labs(color="Origin")

tar_read(fit.systori.strat100)$trajectory %>% 
  mutate(system=factor(system,levels=c("forest","old_sf","young_sf","plantation","agroforestry")),
         origin=factor(origin,levels=c("planted","recruited","remnant"))) %>% 
  ggplot()+
  geom_point(data=data.strat.100,aes(dbh,H,color=origin),alpha=0.1,guide='none')+
  geom_ribbon(aes(dbh,ymin=H_q05,ymax=H_q95,group=origin),fill="grey",alpha=0.5)+
  geom_line(aes(dbh,H_med,group=origin,color=origin))+
  facet_wrap(~system,nrow=1)+
  scale_color_manual(labels=c("Planted","Recruited","Remnant"),
                     values=c("darkred","steelblue2","forestgreen"))+
  scale_y_log10()+
  theme_light()+
  theme(legend.position = "top")+
  labs(color="Origin")

tar_read(fit.systori.strat50)$trajectory %>% 
  mutate(system=factor(system,levels=c("forest","old_sf","young_sf","plantation","agroforestry")),
         origin=factor(origin,levels=c("planted","recruited","remnant"))) %>% 
  ggplot()+
  geom_point(data=data.strat.50,aes(dbh,H,color=origin),alpha=0.1,guide='none')+
  geom_ribbon(aes(dbh,ymin=H_q05,ymax=H_q95,group=origin),fill="grey",alpha=0.5)+
  geom_line(aes(dbh,H_med,group=origin,color=origin))+
  facet_wrap(~system,nrow=1)+
  scale_color_manual(labels=c("Planted","Recruited","Remnant"),
                     values=c("darkred","steelblue2","forestgreen"))+
  scale_y_log10()+
  theme_light()+
  theme(legend.position = "top")+
  labs(color="Origin")
```



# Parameters of the "systori" model

```{r}
fit<-tar_read(global_fit_systori)$fit_df

corr_tab=correspondance_table %>% 
  mutate(system=case_when(system=="young_sf"~"Young sec. for.",
                          system=="old_sf"~"Old sec. for.",
                          TRUE~str_to_title(system)))
fit %>% 
  select(matches(c("alpha","beta"))) %>% 
  pivot_longer(everything()) %>%
  group_by(name) %>%
  summarise(
    q05 = quantile(value, 0.05, na.rm = TRUE),
    q25 = quantile(value, 0.25, na.rm = TRUE),
    q50 = quantile(value, 0.50, na.rm = TRUE),
    q75 = quantile(value, 0.75, na.rm = TRUE),
    q95 = quantile(value, 0.95, na.rm = TRUE)
  ) %>% 
  separate(name,c("parameter", "system", "origin"), convert = TRUE) %>% 
  mutate(corr=paste(system,origin,sep="_")) %>% 
  select(-origin,-system) %>% 
  left_join(corr_tab[,c(3:5)],by="corr") %>% 
  pivot_wider(names_from = parameter,
              values_from = c("q05","q25","q50","q75","q95"), 
              names_glue = "{parameter}_{.value}") %>% 
  ggplot(aes(x=beta_q50, y=alpha_q50,
             col = paste0(system, "\n", origin),
             label = paste0(system, "\n", origin))) +
  geom_segment(aes(x = beta_q25, xend = beta_q75, yend=alpha_q50))+ 
  geom_segment(aes(x = beta_q05, xend = beta_q95, yend=alpha_q50), size = .2) +
  geom_segment(aes(y = alpha_q25, yend = alpha_q75,xend=beta_q50)) +
  geom_segment(aes(y = alpha_q50, yend = alpha_q95,xend=beta_q50), size = .2) +
  geom_point()+
  ggrepel::geom_text_repel(box.padding = 0.4, ) +
  theme_bw() +
  scale_color_discrete(guide = "none") +
  ylab("Asymptotic height [ m ]") +
  xlab("Diameter at 50% height [ cm ]")
```


```{r eval=FALSE, include=FALSE}
fit %>% 
  select(matches(c("alpha","beta"))) %>% 
  pivot_longer(everything()) %>%
  separate(name,c("parameter", "system", "origin"), convert = TRUE) %>% 
  mutate(corr=paste(system,origin,sep="_")) %>% 
  select(-origin,-system) %>% 
  left_join(corr_tab[,c(3:5)],by="corr") %>% 
  ggplot(aes(value,fill=origin))+
  geom_density(alpha=0.5,color=NA)+
  facet_wrap(system~parameter,scales="free")
```


<!-- ## With fit from subdataset -->

```{r eval=FALSE, include=FALSE}
traj_sub<-tar_read(subdata_fit_systori)$trajectory
fit_sub<-tar_read(subdata_fit_systori)$fit_df
tar_load(mod.data)
tar_load(correspondance_table)

trajectories<-ggplot(data=traj_sub)+
  geom_point(data=mod.data,aes(dbh,H,color=system),alpha=0.1)+
  geom_ribbon(aes(dbh,ymin=H_q05,ymax=H_q95,group=system),fill="grey",alpha=0.5)+
  geom_line(aes(dbh,H_med,group=system,color=system))+
  facet_wrap(~origin)+
  scale_color_manual(labels=c("Forest","Secondary forest","Plantation","Agroforestry"),
                     values=c("forestgreen","lightgreen","steelblue3","lightsalmon"))+
  scale_y_log10()+
  theme(legend.position = "top")+
  labs(color="System")
alpha<-fit_sub |> 
  select(matches("alpha")&!matches("_0")) |> 
  pivot_longer(cols=everything()) |> 
  mutate(name=str_replace(name,"alpha_",""))|>
  left_join(correspondance_table,by=c("name"="corr")) |> 
  mutate(system=ordered(system,
                        levels=c("forest","secondary_forest","plantation","agroforestry"))) |> 
  ggplot(aes(value,system,fill=system))+
  geom_density_ridges(scale=0.8)+
  scale_fill_manual(labels=c("Forest","Secondary forest","Plantation","Agroforestry"),
                     values=c("forestgreen","lightgreen","steelblue3","lightsalmon"))+
  facet_wrap(~origin)+
  theme(axis.text.y = element_text(color="white",angle=90),
        axis.ticks.y = element_line(color="white"))+
  theme(legend.position = "none")+
  labs(y="",x="Alpha value")
beta<-fit_sub |> 
  select(matches("beta")&!matches("_0")) |> 
  pivot_longer(cols=everything()) |> 
  mutate(name=str_replace(name,"beta_","")) |> 
  left_join(correspondance_table,by=c("name"="corr")) |> 
  mutate(system=ordered(system,
                        levels=c("forest","secondary_forest","plantation","agroforestry"))) |>
  ggplot(aes(value,system,fill=system))+
  geom_density_ridges(scale=0.8)+
  scale_fill_manual(labels=c("Forest","Secondary forest","Plantation","Agroforestry"),
                     values=c("forestgreen","lightgreen","steelblue3","lightsalmon"))+
  facet_wrap(~origin)+
  theme(axis.text.y = element_text(color="white",angle=90),
        axis.ticks.y = element_line(color="white"))+
  theme(legend.position = "none")+
  labs(y="",x="Beta value")
cowplot::plot_grid(trajectories,alpha,beta,nrow=3,rel_heights = c(1,0.6,0.6))
```

# Biomass predictions

```{r}
tar_load(predict_spatial_crossval)
dataWD <- getWoodDensity(
  genus = predict_spatial_crossval$genusCorr,
  species = predict_spatial_crossval$speciesCorr,
  stand = predict_spatial_crossval$id_plot)
predict_spatial_crossval$AGBobs<- computeAGB(D = predict_spatial_crossval$dbh,
                                             WD = dataWD$meanWD,
                                             H = predict_spatial_crossval$H)
predict_spatial_crossval$AGBpred<- computeAGB(D = predict_spatial_crossval$dbh,
                                              WD = dataWD$meanWD,
                                              H = predict_spatial_crossval$H_med)
Hpred_chave <- computeAGB(D = predict_spatial_crossval$dbh,
                                              WD = dataWD$meanWD,
                                              H = predict_spatial_crossval$H_med)
# rbind(predict_spatial_crossval,
#       predict_fit_systori) |>
```



## Comparison of height prediction between null model and systori model

```{r,fig.height=7,fig.width=10}
predict_spatial_crossval |> 
  mutate(dh=100*(H_med-H)/H,
         sep_h=H-H_med,
         dAGB=100*(AGBpred-AGBobs)/AGBobs,
         mod=factor(mod,levels=c("nul_mod","origin","system","systori_sub","systori","complete"))) |> 
  filter(mod%in%c("nul_mod","systori")) |> 
  ggplot(aes(sep_h,fill=as.factor(mod)))+
  geom_density(alpha=0.5,col=NA)+
  geom_vline(xintercept=0,color="black")+
  theme_light()+
  theme(legend.position = "bottom")+
  facet_wrap(~system,scales="free")+
  labs(fill="Model",
       y="Density",
       x="Error in predicted height (meter)")
```

```{r,eval=FALSE, include=FALSE,fig.height=5,fig.width=10}
predict_spatial_crossval |> 
  mutate(dh=100*(H_med-H)/H,
         sep_h=H-H_med,
         dAGB=100*(AGBpred-AGBobs)/AGBobs,
         mod=factor(mod,levels=c("nul_mod","origin","system","systori_sub","systori","complete"))) |> 
  filter(mod%in%c("nul_mod","systori")) |> 
  ggplot(aes(sep_h,fill=as.factor(mod)))+
  geom_density(alpha=0.5,col=NA)+
  geom_vline(xintercept=0,color="black")+
  theme_light()+
  theme(legend.position = "bottom")+
  facet_wrap(~origin,scales="free")+
  labs(fill="Model",
       y="Density",
       x="Error in predicted height (meter)")
```

## Comparison of biomass prediction error, by tree and between four models
```{r,fig.height=5,fig.width=7}
predict_spatial_crossval |> 
mutate(dh=100*(H_med-H)/H,
         sep_h=H-H_med,
         dAGB=100*(AGBpred-AGBobs)/AGBobs,
         mod=factor(mod,levels=c("nul_mod","origin","system","systori_sub","systori","complete"))) |> 
  filter(mod!="complete") |> 
  ggplot(aes(dAGB,as.factor(mod),fill=as.factor(mod)))+
  geom_boxplot()+
  geom_vline(xintercept=0)+
  theme_light()+
  theme(legend.position="none")+
  facet_wrap(~system)+
  labs(y="Model",
       x="Error in predicted biomass (%)")+
    xlim(c(-100,200))
```




##  Comparison of biomass prediction error, by plot and between four models

```{r,fig.height=5,fig.width=7}
predict_spatial_crossval |> 
  mutate(dh=(H_med-H)/H,
         sep_h=H-H_med,
         dAGB=100*(AGBpred-H_med)/AGBobs,
         mod=factor(mod,levels=c("nul_mod","origin","system","systori_sub","systori","complete"))) |> 
  filter(mod!="complete") |> 
  group_by(system,id_plot,mod) |> 
  summarise(AGBpred_plot=sum(AGBpred),
            AGBobs_plot=sum(AGBobs),
            dAGB_plot=100*(AGBpred_plot-AGBobs_plot)/AGBobs_plot) |> 
  ggplot(aes(dAGB_plot,as.factor(mod),fill=as.factor(mod)))+
  geom_boxplot()+
  geom_vline(xintercept=0)+
  theme_light()+
  theme(legend.position="none")+
  facet_wrap(~system)+
  labs(y="Model",
       x="Error in predicted biomass per plot (%)")
```

<!-- ## from subdata -->

```{r eval=FALSE, include=FALSE}
predict_fit_systori<-tar_read(subdata_fit_systori)$data_pred |>
  select(!c(cof,num,corr)) |> mutate(mod="systori_sub",cluster=as.factor(0))
dataWD <- getWoodDensity(
  genus = predict_fit_systori$genusCorr,
  species = predict_fit_systori$speciesCorr,
  stand = predict_fit_systori$id_plot)
predict_fit_systori$AGBobs<- computeAGB(D = predict_fit_systori$dbh,
                                             WD = dataWD$meanWD,
                                             H = predict_fit_systori$H)
predict_fit_systori$AGBpred<- computeAGB(D = predict_fit_systori$dbh,
                                              WD = dataWD$meanWD,
                                              H = predict_fit_systori$H_med)
# Hpred_chave <- computeAGB(D = predict_fit_systori$dbh,
                                              # WD = dataWD$meanWD,
                                              # H = predict_fit_systori$H_med)

```


<!-- # block map -->
```{r eval=FALSE, include=FALSE}
civ_map=sf::read_sf(dsn="data/civ/",
                layer="COTE D'IVE")

predict_spatial_crossval |>
  select(systori,system,id_plot,lat,long,cluster) |>
  distinct() |>
  ggplot(aes(lat,long,color=cluster))+
  geom_point() +
  facet_wrap(~system)

```

